Draft genomes of the three northern hemisphere blue mussel lineages: North and South European \emph{Mytilus edulis} and Mediterranean \emph{Mytilus galloprovincialis}.

### 2022-06-17 manuscript/ms_genomes.tex
additions

Therefore, need for diverse references among the complex...

@@

We used the program \texttt{AGOUTI} (\url{https://github.com/svm-zhang/AGOUTI}; commit a7e65d6; \cite{Zhang2016a}) to improve scaffolding by using paired end RNA-seq reads.
For each species, a different set of published transcriptomes were used (Supplementary Table [[...]] for accessions).
\texttt{AGOUTI} require a gene prediction as input in addition to RNA-seq reads.
We used \texttt{Augustus} (v3.3.3; \cite{Stanke2008}) to produce an intermediate annotation for each assembly using the \textit{Caenorhabditis} model.
RNA-seq reads were first cleaned using \texttt{rcorrector} (v1.0.4; \cite{Song2015}) and \texttt{trimgalore} (v0.6.6; \url{https://www.bioinformatics.babraham.ac.uk/projects/trim_galore}; \texttt{--quality 20 --stringency 1 -e 0.1 --length 70}).
RNA-seq reads were mapped independently using \texttt{bwa-mem2} (v2.2.1; \cite{VasimuddinMd2019}), and then merged as a common bam file for each assembly with \texttt{samtools} (v1.12; \cite{Li2009}).
Finally, the \texttt{AGOUTI} scaffolding pipeline was run using the gene prediction and mapped RNA-seq reads (-minMQ 20 -maxFracMM 0.05; using \texttt{python} v2.7.15 and \texttt{samtools} v1.10 for compatibility).

@@

Finally, gene completion analyses were carried out using \texttt{BUSCO} (v5.1.1; \cite{Manni2021}).
We a Metazoan (\texttt{metazoa\_odb10.2021-02-24}) and a Molluscan database (\texttt{mollusca\_odb10.2020-08-05}) to assess and compare new and published assemblies.
Published assemblies are:
\begin{itemize}
	\item \textit{M. coruscus}, GCA\_017311375.1, \textcite{Yang2021};
	\item \textit{M. galloprovincialis} from the Altantic lineage, GCA\_900618805.1, \textcite{Gerdol2020};
	\item \textit{M. edulis} from the Southern European lineage, GCA\_905397895.1, \textcite{Corrochano-Fraile2021}
	\item \textit{M. edulis} from the American lineage, GCA\_019925275.1.
\end{itemize}

@@

\subsection{Assembly statistics}

@@

\subsection{BUSCO scores}

@@

\section{Conclusion}\label{conclusion}

\section{Acknowledgement}

@@

% Add supplementary here
\appendix

>>>>

As an effort to diversify the genomic resources available for the \textit{Mytilus} species complex,
we assembled and annotate the genomes of three lineages using the 10X chromium technology.
While our assemblies were initially fragmented due to the high level of heterozygosity,
we leveraged the existence of a chromosome scale assembly of a sister species to scaffold our assemblies against it.
We obtained assemblies equivalent to published ones in term of completeness for three new lineages of the species complex using a low sequencing budget and publicly available data.
The resources produced and the assembly pipeline are freely available for use by the community.

@@

We used \texttt{AGOUTI} (\url{https://github.com/svm-zhang/AGOUTI}; commit a7e65d6; \cite{Zhang2016a}) to improve scaffolding by using paired end RNA-seq reads.
For each species, a different set of published transcriptomes were used (Supplementary Table [[...]] for accessions).
\texttt{AGOUTI} require a gene prediction as input in addition to RNA-seq reads.
We used \texttt{Augustus} (v3.3.3; \cite{Stanke2008}) to produce an intermediate annotation for each assembly using the \textit{Caenorhabditis} model.
RNA-seq reads were first cleaned using \texttt{rcorrector} (v1.0.4; \cite{Song2015}) and \texttt{trimgalore} (v0.6.6; \url{https://www.bioinformatics.babraham.ac.uk/projects/trim_galore}; \texttt{--quality 20 --stringency 1 -e 0.1 --length 70}).
RNA-seq reads were mapped independently using \texttt{bwa-mem2} (v2.2.1; \cite{VasimuddinMd2019}), and then merged as a common bam file for each assembly with \texttt{samtools} (v1.12; \cite{Li2009}).
Finally, the \texttt{AGOUTI} scaffolding pipeline was run using the gene prediction and mapped RNA-seq reads (-minMQ 20 -maxFracMM 0.05; using \texttt{python} v2.7.15 and \texttt{samtools} v1.10 for compatibility).

@@

Finally, gene completion analyses were carried out using \texttt{BUSCO} (v5.1.1; \cite{Manni2021}).
We used both a Metazoan (\texttt{metazoa\_odb10.2021-02-24}) and Molluscan (\texttt{mollusca\_odb10.2020-08-05}) database to assess and compare new and published assemblies.
We compared our assemblies to the following published ones:
\begin{itemize}
	\item \textit{M. coruscus}, GCA\_017311375.1, \textcite{Yang2021};
	\item \textit{M. galloprovincialis} from the Altantic lineage, GCA\_900618805.1, \textcite{Gerdol2020};
	\item \textit{M. edulis} from the Southern European lineage, GCA\_905397895.1, \textcite{Corrochano-Fraile2021}
	\item \textit{M. edulis} from the American lineage, GCA\_019925275.1.
\end{itemize}

@@

The pre-assembly k-mer analysis carried out using Genomescope (using 21-mers) showed that, as expected, the genomes were highly heterozygous with values going from 3.49 to 8.77~\% (\cref{fig:genomescope}).
Genomescope provide a bad fit and estimation in the case of the MeduEUS data due to a reduced coverage of the genome in that case, compared to the two other assemblies.

Highly heterozygous genomes brings the risk of having a large number of duplicate contigs due to the separate assembly of the two alleles from a same locus.
For this reason, we used the \texttt{purge\_dups} pipeline to try removing a maximum of such bias.
This procedure reduced the number of complete duplicated genes in all assemblies (\cref{supfig:busco}, v1 to v4).

\begin{figure}[h]
	\includegraphics[width=\linewidth]{figures/Fig1_genomescope.pdf}
	\caption{Genomescope}
	\label{fig:genomescope}
\end{figure}

@@

To assess the completeness of our assemblies, we compared them to four \textit{Mytilus} published assemblies on the basis of a Metazoan and a Molluscan set of single copy orthologous genes using BUSCO (\cref{fig:busco}).
Overall, we show that our assemblies are equivalent to the published ones in terms of completeness.

@@

%\section{Conclusion}\label{conclusion}

\section*{Acknowledgements}

@@

% Add supplementary here
\newpage
\appendix
\setcounter{table}{0}
\renewcommand{\thetable}{S\arabic{table}}
\setcounter{figure}{0}
\renewcommand{\thefigure}{S\arabic{figure}}
\section*{Supplementary Information}

\begin{figure}[h]
	\includegraphics[width=\linewidth]{figures/supfig_busco.pdf}
	\caption{Busco scores}
	\label{supfig:busco}
\end{figure}
